---
title: "starting a new project like a boss"
author: "Rose Hartman"
date: "June 30, 2016"
output: html_document
---

1. Pick a name for the project, for example if you're studying awesomeness, you might want to call it "awesomeness-study".

2. Navigate to the directory where you want this new project folder to be created. 
I have a folder called STUDIES where I keep all of my project directories. 

```{r}
setwd("/Users/TARDIS/Documents/STUDIES")
```

3. Use ProjectTemplate to automatically generate a bunch of useful stuff.

```{r}
library(ProjectTemplate)
create.project("awesomeness-study")
```

This makes a new directory called awesomeness-study and a bunch of directories and files within it to get you started with a lovely, organized code project.

4. Make that an RStudio project. File > New Project > Associate a project with an existing working directory.

5. Turn on version control. Tools > Project Options > Git/SVN > version control system: git

6. Go to your github and make a new repo, probably named awesomeness-study. 

7. When you make the repo, you'll see directions to push an existing repository from the command line:
```{r eval=FALSE}
git remote add origin https://github.com/rosemm/awesomeness-study.git
git push -u origin master
```
This should be run in Terminal. You can open Terminal easily in your current working directory from RStudio: Tools > Shell. Then just run those two lines of code.

